Enhancing Multiple Myeloma Cell Segmentation through Physical Image Transformations

### 2023-05-02 paper/paper.tex
Update paper

In Gupta A et al [3], the paper discussed multiple approaches to tackle the challenge on segmentation of Multiple Myeloma cells on microscopic images. The top solution combined three instance segmentation models with three convolutional feature extractors, extensive image augmentation, and a custom model ensemble strategy. It utilized a post-processing step to pair corresponding instances of nuclei and cytoplasm together, using an auxiliary whole-cell class. The second-placed solution used a two-stage approach where the first stage consisted of extraction of possible nucleus instances using a U-Net and the second stage relied on dilated convolution and attention mechanism using Attention Deeplabv3+ for cytoplasm segmentation.

@@

\section{Methodology}
\subsection{Data Sources \& Processing}
For this work, we used the SegPC-2021 challenge dataset, a medical imaging challenge on on "Segmentation of Multiple Myeloma Plasma Cells in Microscopic Images (SegPC-2021)" organized at the IEEE International Symposium on Biomedical Imaging (ISBI), 2021, France. The dataset was obtained from patients diagnosed or treated for Multiple Myeloma (MM) at AIIMS, New Delhi, India. Microscopic images were captured from bone marrow aspirate slides using two cameras: 1. Olympus camera with an image size of 2040x1536 pixels and 2. Nikon camera with an image size of 1920x2560 pixels. The images were captured using a stain color normalization technique to reduce the color variations in the images and were divided into three sets:
\begin{enumerate}
  \item Train (298 images)
  \item Validation (200 images)
  \item Test (277 images)
\end{enumerate}
The Ground Truth (GT) values were provided for the train and validation sets, but not for the test set. These were provided in form of a mask where each mask is an image with a single segmented Multiple Myeloma cell. The background of the mask had a pixel value of 0, the nucleus of the cell was represented by a pixel value of 40, and the cytoplasm of the cell was represented by a pixel value of 20. The masks were of the same size as the original images.

@@

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=.4\linewidth]{1.png}
    \captionof{figure}{Original Image}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \centering
    \includegraphics[width=.6\linewidth]{2.jpg}
    \captionof{figure}{Simulation of 12 light waves hitting a sample from different angles
    }
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=.4\linewidth]{3.png}
    \captionof{figure}{Image with contrast normalization}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \centering
    \includegraphics[width=.6\linewidth]{4.png}
    \captionof{figure}{Image after Dilation operation}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=.4\linewidth]{5.png}
    \captionof{figure}{Image after Erosion operation}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \centering
    \includegraphics[width=.6\linewidth]{6.png}
    \captionof{figure}{Image with Sobel Gradient filter}
  \end{minipage}
\end{figure}
\subsubsection{Blur Filters}

@@

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{7.png}
    \captionof{figure}{Image after Erosion operation}
  \end{minipage}%
\end{figure}

@@

@@

Multiple Myeloma is a challenging and complex type of blood cancer that requires accurate identification and segmentation of plasma cells in order to aid in diagnosis, prognosis, and treatment planning. In this study, we have explored the effectiveness of various physical image transformations to improve the accuracy of Multiple Myeloma cell segmentation. Our results have shown that certain physical image transformations, such as Dilation and Median Blur filter, were effective in improving the Average Precision and Mean IOU of our baseline model. These transformations helped to enhance the contrast and reduce noise in the images, resulting in more accurate and consistent segmentation of Multiple Myeloma cells.

>>>>

In Gupta A et al [3], the paper discussed multiple approaches to tackle the challenge of segmentation of Multiple Myeloma cells on microscopic images. The top solution combined three instance segmentation models with three convolutional feature extractors, extensive image augmentation, and a custom model ensemble strategy. It utilized a post-processing step to pair corresponding instances of nuclei and cytoplasm together, using an auxiliary whole-cell class. The second-placed solution used a two-stage approach where the first stage consisted of extraction of possible nucleus instances using a U-Net and the second stage relied on dilated convolution and attention mechanism using Attention Deeplabv3+ for cytoplasm segmentation.

@@

\section{Methodology}
\subsection{Data Sources \& Processing}
For this work, we used the SegPC-2021 challenge dataset, a medical imaging challenge on the "Segmentation of Multiple Myeloma Plasma Cells in Microscopic Images (SegPC-2021)" organized at the IEEE International Symposium on Biomedical Imaging (ISBI), 2021, France. The dataset was obtained from patients diagnosed or treated for Multiple Myeloma (MM) at AIIMS, New Delhi, India. Microscopic images were captured from bone marrow aspirate slides using two cameras: 1. Olympus camera with an image size of 2040x1536 pixels and 2. Nikon camera with an image size of 1920x2560 pixels. The images were captured using a stain color normalization technique to reduce the color variations in the images and were divided into three sets:
\begin{enumerate}
  \item Train (298 images)
  \item Validation (200 images)
  \item Test (277 images)
\end{enumerate}
The Ground Truth (GT) values were provided for the train and validation sets, but not for the test set. These were provided in form of a mask where each mask is an image with a single segmented Multiple Myeloma cell. The background of the mask had a pixel value of 0, the nucleus of the cell was represented by a pixel value of 40, and the cytoplasm of the cell was represented by a pixel value of 20. The masks were of the same size as the original images.

@@

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{1.png}
    \captionof{figure}{Original Image}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \centering
    \includegraphics[width=.8\linewidth]{2.jpg}
    \captionof{figure}{Simulation of 12 light waves hitting a sample from different angles
    }
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{3.png}
    \captionof{figure}{Image after contrast normalization}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \centering
    \includegraphics[width=.63\linewidth]{4.png}
    \captionof{figure}{Image after Dilation operation}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{5.png}
    \captionof{figure}{Image after Erosion operation}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \centering
    \includegraphics[width=.63\linewidth]{6.png}
    \captionof{figure}{Image after Sobel filter}
  \end{minipage}
\end{figure}
\subsubsection{Blur Filters}

@@

\begin{figure}
  \centering
  \begin{minipage}{1\textwidth}
    \centering
    \includegraphics[width=0.8\linewidth]{7.png}
    \captionof{figure}{Image after applying color channel enhancement}
  \end{minipage}%
\end{figure}

\begin{figure}
  \centering
  \begin{minipage}{.5\textwidth}
    \centering
    \includegraphics[width=.5\linewidth]{8.png}
    \captionof{figure}{Image after Gaussian Blur}
  \end{minipage}%
  \begin{minipage}{.4\textwidth}
    \centering
    \includegraphics[width=.63\linewidth]{9.png}
    \captionof{figure}{Image after Median Blur}
  \end{minipage}
\end{figure}

@@

\begin{figure}
  \centering
  \begin{minipage}{1\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{10.png}
    \captionof{figure}{Machine Learning Pipeline \emph{(*trainable layer exists only in illumination simulation case)}}
  \end{minipage}%
\end{figure}

@@

Multiple Myeloma is a challenging and complex type of blood cancer that requires accurate identification and segmentation of plasma cells in order to aid in diagnosis, prognosis, and treatment planning. In this study, we have explored the effectiveness of various physical image transformations to improve the accuracy of Multiple Myeloma cell segmentation. Our results have shown that certain physical image transformations, such as the Dilation and Median Blur filters, were effective in improving the Average Precision and Mean IOU of our baseline model. These transformations helped to enhance the contrast and reduce noise in the images, resulting in more accurate and consistent segmentation of Multiple Myeloma cells.
